A SIMPLE TEXT ANALYTICS MODEL TO ASSIST LITERARY CRITICISM: COMPARATIVE APPROACH AND EXAMPLE ON JAMES JOYCE AGAINST SHAKESPEARE AND THE BIBLE

### 2017-08-24 latex/article.tex
notas e detalhe no titulo
incorporada revisao do luis

%%-----------------------------------------------EDIT-----------------------------------------------
\title{A SIMPLE MODEL FOR LITERARY ANALYSIS BY TEXT ANALYTICS: COMPARATIVE APPROACH AND EXAMPLE ON JAMES JOYCE AGAINST SHAKESPEARE AND THE BIBLE} 
% Mining James Joyce against reference literature: differences, similarities and the ubiquity of the reader

%%-----------------------------------------------EDIT----------------------------------------------
\author
    {\rm \begin{tabular}{l} 
    \textbf{Renato Fabbri}$^{1}$ - {\textnormal renato.fabbri@gmail.com}\\%
    \textbf{Luis Henrique Garcia Ferreira}$^{2}$ - {\textnormal henrique@hotmail.com}\\
    {\fontsize{11}{0}\selectfont $^{1}$University of São Paulo, Institute of Mathematical and Computer Sciences - São Carlos, SP, Brazil}\vspace*{-0.05cm} \\
    {\fontsize{11}{0}\selectfont $^{2}$University of Campinas, Institute of Language Studies - Campinas, SP, Brazil}\vspace*{-0.05cm}\\
  \end{tabular}}
%%----------------------------------------------------------------------------------------------

@@

We should emphasize that changing this reference literature is very straightforward.
One should only provide the corresponding text files and modify the scripts to read the
intended records.
If the works are well-known, the process should require only a quick search on the web
(e.g. within Gutenberg or Archive.org projects), saving the text locally and then changing
filenames in the scripts.
Some possibilities include: other masters of English literature;
a selection of poets; works from scientific literature of a specific field;
works in other language, such as Machado de Assis and Clarice Linspector
if analyzing works in Brazilian Portuguese.
There is no reason why the corpus should not include data streaming (e.g. from Twitter)
or access to online resources,
such as Wikipedia pages.

@@

\begin{figure}[!htbp] %h or !htbp
\vspace{-2pt}
\begin{center}
\includegraphics[height=6.7cm,width=9cm]{figs/pca}%
    \caption{Reference works (from Shakespeare and the Bible) and a selection of books by James Joyce plotted with respect to the first two principal components.
    As expected, Ulysses and Finnegans Wake are near each other and farther from the other books. Dubliners and Portrait are also near each other but also near other books (in this case, books from the Bible).
    Stephen Hero is between these two groups, also coherent with expectations.
    There is no prevalence of few measures in these components, reason why we omit this aspect of the analysis.
    The interested reader should access the scripts described in Table~\ref{tab:files} to deepen the analysis exposed here only by way of illustration of the proposed method.}
\label{fig:pca1}%
\end{center}
\end{figure}

@@

\begin{figure}[!htbp] %h or !htbp
\vspace{-2pt}
\begin{center}
\includegraphics[height=6.7cm,width=9cm]{figs/abst2}%
    \caption{Fraction of known words among all tokens and among most significant words (words which are not stopwords).
    Lower faction is regarded here as evidence of abstraction because the reader should infer meaning.
    Finnegans Wake is very distinct from all the books, as expected.
    It is surprising that: 1) Ulysses has a higher rate of known words that Dubliners;
    and 2) that these two measures are the best for a classifier to identify these works by James Joyce, among all the measures used in the figures of this article, including the principal components.}
\label{fig:abst2}%
\end{center}
\end{figure}

@@

In further efforts, we should:
\begin{itemize}
    \item Deepen the analysis of the reference literature (books by Shakespeare and in the Bible)
        to better contextualize any literature we consider against them.
    \item Expand the use of Wordnet to encompass synonymy, antonymy, meronymy, etc.
        Also to consider specific roots of nouns, adjectives, verbs and adverbs.
    \item Report this endeavor to the literary criticism academic community.
        This should be done at least in two ways: by describing the method and its relevance
        within the humanities background;
        and by exposing results from analyzing specific authors, such as Joyce and Ezra Pound.
    \item Consider other measures of abstraction.
        Can we use length of words and sentences?
        Should we count the root synsets instead of the depth?
    \item Vary the methods and state reasonable generic bounds e.g. for splitting a work to
        obtain more data points.
    \item Investigate the results exposed in Figure~\ref{fig:abst1} which are not in consonance with what we expected.
\end{itemize}

@@

% ------------------------------------------------------------------------
\begin{thebibliography}{99}
\fontsize{11}{0}\selectfont
\bibitem[Boom, 1999]{shake}
Bloom, H. (1998). Shakespeare: The invention of the human. Riverhead Trade (Paperbacks).

>>>>

%%-----------------------------------------------EDIT-----------------------------------------------
\title{A SIMPLE TEXT ANALYTICS MODEL TO ASSIST LITERARY CRITICISM: COMPARATIVE APPROACH AND EXAMPLE ON JAMES JOYCE AGAINST SHAKESPEARE AND THE BIBLE} 
% Mining James Joyce against reference literature: differences, similarities and the ubiquity of the reader

%%-----------------------------------------------EDIT----------------------------------------------
\author
    {\rm \begin{tabular}{l} 
    \textbf{Renato Fabbri}$^{1}$ - {\textnormal renato.fabbri@gmail.com}\\%
    \textbf{Luis Henrique Garcia}$^{2}$ - {\textnormal henriquegarcia.pesquisa@gmail.com}\\
    {\fontsize{11}{0}\selectfont $^{1}$University of São Paulo, Institute of Mathematical and Computer Sciences - São Carlos, SP, Brazil}\vspace*{-0.05cm} \\
    {\fontsize{11}{0}\selectfont $^{2}$University of Campinas, Institute of Language Studies - Campinas, SP, Brazil}\vspace*{-0.05cm}\\
  \end{tabular}}
%%----------------------------------------------------------------------------------------------

@@

We should emphasize that changing this reference literature is very straightforward.
One should only provide the corresponding text files and modify the scripts to read the
intended records.
If the works are well-known, the process should require only a quick search on the web
(e.g. within Gutenberg or Archive.org projects), saving the text locally and then changing
filenames in the scripts.
Some possibilities include: other masters of English literature;
a selection of poets; works from scientific literature of a specific field;
works in other language, such as Machado de Assis and Clarice Lispector
if analyzing works in Brazilian Portuguese.
There is no reason why the corpus should not include data streaming (e.g. from Twitter)
or access to online resources,
such as Wikipedia pages.

@@

\begin{figure}[!htbp] %h or !htbp
\vspace{-2pt}
\begin{center}
\includegraphics[height=6.7cm,width=9cm]{figs/pca}%
    \caption{Reference works (from Shakespeare and the Bible) and a selection of books by James Joyce plotted with respect to the first two principal components.
    As expected, Ulysses and Finnegans Wake are near each other and farther from the other books. Dubliners and Portrait are also near each other but also near other books (in this case, books from the Bible), which is in accordance with their style.
    Stephen Hero is between these two groups, also coherent with expectations.
    There is no prevalence of few measures in these components, reason why we omit this aspect of the analysis.
    The interested reader should access the scripts described in Table~\ref{tab:files} to deepen the analysis exposed here only by way of illustration of the proposed method.}
\label{fig:pca1}%
\end{center}
\end{figure}

@@

\begin{figure}[!htbp] %h or !htbp
\vspace{-2pt}
\begin{center}
\includegraphics[height=6.7cm,width=9cm]{figs/abst2}%
    \caption{Fraction of known words among all tokens and among most significant words (words which are not stopwords).
    Lower fraction is regarded here as evidence of abstraction because the reader should infer meaning.
    Finnegans Wake is very distinct from all the books, as expected.
    It is surprising that: 1) Ulysses has a higher rate of known words that Dubliners;
    and 2) that these two measures are the best for a classifier to identify these works by James Joyce, among all the measures used in the figures of this article, including the principal components.}
\label{fig:abst2}%
\end{center}
\end{figure}

@@

In further efforts, we should:
\begin{itemize}
    \item Deepen the analysis of the reference literature (books by Shakespeare and in the Bible)
        to better contextualize any literature we consider against them.
    \item Expand the use of Wordnet to encompass synonymy, antonymy, meronymy, etc.
        Also to consider specific roots of nouns, adjectives, verbs and adverbs.
    \item Report this endeavor to the literary criticism academic community.
        This should be done at least in two ways: by describing the method and its relevance
        within the humanities background;
        and by exposing results from analyzing specific authors, such as Joyce and Ezra Pound.
    \item Consider other measures of abstraction.
	    Should we regard the length of words and sentences as cues of an author's style?
        Should we count the root synsets instead of the depth?
    \item Vary the methods and state reasonable generic bounds e.g. for splitting a work to
        obtain more data points.
    \item Investigate the results exposed in Figure~\ref{fig:abst1} which are not in consonance with what we expected.
\end{itemize}

@@

% ------------------------------------------------------------------------
\begin{thebibliography}{99}
\fontsize{11}{0}\selectfont
\bibitem[Bloom, 1998]{shake}
Bloom, H. (1998). Shakespeare: The invention of the human. Riverhead Trade (Paperbacks).
